Model Based Systems Engineering with a Docs-as-Code Approach for the SeaLion CubeSat Project

### 2023-04-24 manuscript/template.tex
Incorporate peer comments. Corrected references and citations and resized figures.

\subsection{Documents-As-Code Approach}
Documents-as-code (Docs-as-code) refers to a philosophy that team members should be writing documentation with the same tools as code \cite{ibm_mbse}.  This allows for documentation to updated seamlessly without additional work with document tools (doctools).  The code tools would include version control (e.g., Git), issues trackers, code tools (e.g., Visual Studio Code), etc.  To do so would mean that the “following the same workflows as development teams, and being integrated in the product team. It enables a culture where writers and developers both feel ownership of documentation, and work together to make it as good as possible” \cite{ibm_mbse}.  The SeaLion team taking advantage of the aforementioned philosophy would realize the benefits of utilizing the same principles and practices used to manage software, using modern version control tools (e.g., Git), for the configuration management of mission and flight software architecture documentation, and captured in a model-based approach \cite{ibm_mbse}.  In addition, these models can stored and used persistently on a local file system without the use of cloud based services.  Proprietary services are not required to generate documentation, modify documentation, or modify models.  Similar approaches have been seen as Structurizr \cite{structurizr} and F Prime Prime (FPP) \cite{f_prime_prime}.  FPP is based on F Prime which is an open-source software framework developed by NASA's Jet Propulsion Laboratory \cite{f_prime}.  While docs-as-code has precedent, the methodologies noted are not easily accesible to those without a sufficient background in programming.

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/user_story.png}
    \caption{Example YAML file}
	\label{fig:example_yaml}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/reference.png}
    \caption{References YAML file}
	\label{fig:reference_yaml}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/uml_stakeholder.png}
    \caption{Stakeholder UML Diagram}
	\label{fig:uml_stakeholder}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/uml_userstory.png}
    \caption{UML diagram of user stories}
	\label{fig:uml_userstory}
    \end{figure}   
\unskip

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/uml_userstory2.png}
    \caption{UML diagram of ground station operator}
	\label{fig:uml_userstory2}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/datastructure.png}
    \caption{Satellite Health Data Structure}
	\label{fig:datastructure}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/uml_datastructure.png}
    \caption{UML Diagram of Data Structures}
	\label{fig:uml_datastructure}
    \end{figure}   
\unskip

\subsection{Document Generation}
As noted multiple times throughout this article, there have been a number of figures and tables generated from the YAML files placed within the SeaLion mission architecture GitHub repository.  Many of the figures are UML diagrams that are auto-generated artifacts rendered from the M30ML modeling language and formatted using the Liquid template language.  This is the link for the docs-as-code approach.  YAML code files are used to generate documents for information sharing between group members.  This means that any changes made to the SeaLion mission architecture model can immediately be used to generate new documents.  Whether it be diagrams, tables, or text, continuous updating is ensured that any changes affecting dependencies within the mission architecture are kept in sync.  YAML files are automatically processed using templates langauge (e.g., Jinja2) via a build shell script.  This also means that content of the model and formatting of documents are decoupled.  The model is formatting-agnostic for documentation purposes as this is handled by template language.  A conference proceeding manuscript presented in AIAA SciTech 2023 was created purely by a docs-as-code format \cite{scitech_proceeding}.  The team used a LaTeX template to automatically format the manuscript to the conference guidelines and subsequently inject items such as the generated diagrams, tables, and references directly into the manuscript.

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/components.png}
    \caption{Component YAML File}
	\label{fig:components}
    \end{figure}   
\unskip

@@

\acknowledgments{Special thanks to Dr. Sharan Asundi, PhD., Sean Marquez, and the SeaLion Project Team}

@@

\noindent 
\begin{tabular}{@{}ll}
1U &	1-Unit\\
2U &	2-Unit\\
3U &	3-Unit\\
AC &	Alternating Current\\
AFIT &	 Air Force Institute of Technology \\
AIAA &	American Institute of Aeronautics and Astronautics \\
AODS &	Altitude and Orbit Determination System\\
BOM &	Bill of Material\\
ConOps &	Concept of Operations\\
COTS &	Commerical Off-the-Shelf\\
DeCS &	Deployable Composite Structure\\
DOA &	Dead on Arrival\\
doctools &	Dcoument Tools\\
DOF &	Distributed OSHW Framework \\
EVR &	Event\\
GPS &	Global Positioning System\\
ISS &	International Space Station\\
M30ML &	Mach 30 Modelling Language\\
MBSE &	Model Based Systems Engineering\\
MC3 &	Mobile CubeSat Command and Control \\
Me-S &	Multi-spectral Sensor \\
NRL &	Naval Research Laboratory\\
ODU &	Old Dominion University\\
OML &	Ontological Modeling Language \\
OSHW &	Open Source Hardware\\
OWL2 &	Web Ontology Language 2 \\
Q4 &	Quarter Four\\
SPADE &	Space PlasmADiagnostic suitE\\
SWRL &	Semantic Web Rule Language \\
UHF &	Ultra High Frequency \\
UML & Unified Modelling Language\\
USCGA &	United States Coast Guard Academy\\
VLEO &	Very Low Earth Orbit\\
WFF &	Wallops Flight Facility 
\end{tabular}
}

>>>>

\subsection{Documents-As-Code Approach}
Documents-as-code (Docs-as-code) refers to a philosophy that team members should be writing documentation with the same tools as code \cite{docs_as_code}.  This allows for documentation to updated seamlessly without additional work with document tools (doctools).  The code tools would include version control (e.g., Git), issues trackers, code tools (e.g., Visual Studio Code), etc.  To do so would mean that the “following the same workflows as development teams, and being integrated in the product team. It enables a culture where writers and developers both feel ownership of documentation, and work together to make it as good as possible” \cite{docs_as_code}.  The SeaLion team taking advantage of the aforementioned philosophy would realize the benefits of utilizing the same principles and practices used to manage software, using modern version control tools (e.g., Git), for the configuration management of mission and flight software architecture documentation, and captured in a model-based approach \cite{docs_as_code}.  In addition, these models can stored and used persistently on a local file system without the use of cloud based services.  Proprietary services are not required to generate documentation, modify documentation, or modify models.  Similar approaches have been seen as Structurizr \cite{structurizr} and F Prime Prime (FPP) \cite{f_prime_prime}.  FPP is based on F Prime which is an open-source software framework developed by NASA's Jet Propulsion Laboratory \cite{f_prime}; it is a model-driven approach for producing flight software code that can be compiled directly onto flight hardware.  While docs-as-code has precedent, the methodologies noted are not easily accesible to those without a sufficient background in programming.

@@

\begin{figure}[H]
    \includegraphics[width=12 cm]{assets/user_story.png}
    \caption{Example YAML file}
	\label{fig:example_yaml}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=10 cm]{assets/reference.png}
    \caption{References YAML file}
	\label{fig:reference_yaml}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=12 cm]{assets/uml_stakeholder.png}
    \caption{Stakeholder UML Diagram}
	\label{fig:uml_stakeholder}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=12 cm]{assets/uml_userstory.png}
    \caption{UML diagram of user stories}
	\label{fig:uml_userstory}
    \end{figure}   
\unskip

\begin{figure}[H]
    \includegraphics[width=12 cm]{assets/uml_userstory2.png}
    \caption{UML diagram of ground station operator}
	\label{fig:uml_userstory2}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=12 cm]{assets/datastructure.png}
    \caption{Satellite Health Data Structure}
	\label{fig:datastructure}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=12 cm]{assets/uml_datastructure.png}
    \caption{UML Diagram of Data Structures}
	\label{fig:uml_datastructure}
    \end{figure}   
\unskip

\subsection{Document Generation}
As noted multiple times throughout this article, there have been a number of figures and tables generated from the YAML files placed within the SeaLion mission architecture GitHub repository.  Many of the figures are UML diagrams that are auto-generated artifacts rendered from the M30ML modeling language and formatted using the Liquid template language.  This is how the docs-as-code approach is implemented.  YAML code files are used to generate documents for information sharing between group members.  This means that any changes made to the SeaLion mission architecture model can immediately be used to generate new documents.  Whether it be diagrams, tables, or text, continuous updating is ensured that any changes affecting dependencies within the mission architecture are kept in sync.  YAML files are automatically processed using templates langauge (e.g., Jinja2) via a build shell script.  This also means that content of the model and formatting of documents are decoupled.  The model is formatting-agnostic for documentation purposes as this is handled by the template language.  A conference proceeding manuscript presented in AIAA SciTech 2023 was created purely by a docs-as-code format \cite{scitech_proceeding}.  The team used a LaTeX template to automatically format the manuscript to the conference guidelines and subsequently inject items such as the generated diagrams, tables, and references directly into the manuscript.

@@

\begin{figure}[H]
    \includegraphics[width=5 cm]{assets/components.png}
    \caption{Component YAML File}
	\label{fig:components}
    \end{figure}   
\unskip

@@

\acknowledgments{Special thanks to the SeaLion Project Team}

@@

\noindent 
\begin{tabular}{@{}ll}
1U &	1-Unit\\
2U &	2-Unit\\
3U &	3-Unit\\
AC &	Alternating Current\\
AFIT &	 Air Force Institute of Technology \\
AIAA &	American Institute of Aeronautics and Astronautics \\
AODS &	Altitude and Orbit Determination System\\
BOM &	Bill of Material\\
ConOps &	Concept of Operations\\
COTS &	Commerical Off-the-Shelf\\
DeCS &	Deployable Composite Structure\\
DOA &	Dead on Arrival\\
doctools &	Dcoument Tools\\
DOF &	Distributed OSHW Framework \\
EVR &	Event\\
FPP &	F Prime Prime\\
GPS &	Global Positioning System\\
ISS &	International Space Station\\
M30ML &	Mach 30 Modelling Language\\
MBSE &	Model Based Systems Engineering\\
MC3 &	Mobile CubeSat Command and Control \\
Me-S &	Multi-spectral Sensor \\
NRL &	Naval Research Laboratory\\
ODU &	Old Dominion University\\
OML &	Ontological Modeling Language \\
OSHW &	Open Source Hardware\\
OWL2 &	Web Ontology Language 2 \\
Q4 &	Quarter Four\\
SPADE &	Space PlasmADiagnostic suitE\\
SWRL &	Semantic Web Rule Language \\
UHF &	Ultra High Frequency \\
UML & Unified Modelling Language\\
USCGA &	United States Coast Guard Academy\\
VLEO &	Very Low Earth Orbit\\
WFF &	Wallops Flight Facility 
\end{tabular}
}
